### 2020-11-23 Thesis-Chapter2-Background.tex
Spelling and editorial improvements

\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
locked iPhone, resulted in the first major battle of \ac{the-cw2}. The \ac{FBI} issued an order under the \acrlong{AWA}
to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
February 16, 2016, when the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the
phone through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto
wars---terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
% (https://www.chicagomanualofstyle.org/qanda/data/faq/topics/HyphensEnDashesEmDashes/faq0002.html)
jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
make it to the floor, and the unresolved nature of the court case left the legal debate unresolved.

@@

\paragraph*{Volatile Politics} Finally, \ac{the-cw2} is taking place amongst a political landscape marked by
uncertainty, extremism, misinformation, and a pandemic that will have unknown long-term  impacts. Since legislature is
part of the broader data security threat model, its behavior is important. Unfortunately, the dynamics of present
leadership and the lack of reliable data on what law enforcement agencies need make it difficult for the technical
community to engage productively with policy makers \cite{granick_2018}.

@@

Among European countriues, France and Russia join the U.K. in having access mandates \cite{lewis_2017}. France, Belgium,
Germany, and Sweden require cooperation between telecommunications providers and law enforcement; for France, Belgium,
and Germany, that includes decrypting network traffic when possible. France, Germany, and Sweden each have some level of
\ac{lawful-hacking} capability. There is no E.U. legislation that requires key disclosure or decryption of network
traffic.

@@

Some schemes compensate for the risk of concentrated sensitive information by massively distributing the secret
information. In this arrangement, trust is put in the distrubuted system, not in any single party. Such systems offer
high data availability while making it difficult for attempts to acquire keys to be secret or to subvert access
policies. This approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has modern blockchain
\cite{phan_key_2017} and distributed device \cite{servan_schreiber_jje_2020} variations. These approaches are not as
thoroughly explored as trusted-party \ac{key-escrow}.

>>>>

\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
locked iPhone, resulted in the first major battle of \ac{the-cw2}. The \ac{FBI} issued an order under the \acrlong{AWA}
to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
February 16, 2016, when the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the
phone through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto
wars---terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
make it to the floor, and the unresolved nature of the court case left the legal debate unresolved.

@@

\paragraph*{Volatile Politics} Finally, \ac{the-cw2} is taking place amongst a political landscape marked by
uncertainty, extremism, misinformation, and a pandemic that will have unknown long-term impacts. Since legislature is
part of the broader data security threat model, its behavior is important. Unfortunately, the dynamics of present
leadership and the lack of reliable data on what law enforcement agencies need make it difficult for the technical
community to engage productively with policymakers \cite{granick_2018}.

@@

Among European countries, France and Russia join the U.K. in having access mandates \cite{lewis_2017}. France, Belgium,
Germany, and Sweden require cooperation between telecommunications providers and law enforcement; for France, Belgium,
and Germany, that includes decrypting network traffic when possible. France, Germany, and Sweden each have some level of
\ac{lawful-hacking} capability. There is no E.U. legislation that requires key disclosure or decryption of network
traffic.

@@

Some schemes compensate for the risk of concentrated sensitive information by massively distributing the secret
information. In this arrangement, trust is put in the distributed system, not in any single party. Such systems offer
high data availability while making it difficult for attempts to acquire keys to be secret or to subvert access
policies. This approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has modern blockchain
\cite{phan_key_2017} and distributed device \cite{servan_schreiber_jje_2020} variations. These approaches are not as
thoroughly explored as trusted-party \ac{key-escrow}.

### 2020-11-23 Thesis-Chapter3-PolicyMaking.tex
Spelling and editorial improvements

The failures of the \ac{classical-method} and \ac{incrementalism} teach that we must be able to think strategically
while respecting the non-linear nature of \acp{wicked-prob}. One way to use the high level, holistic view of the
classical approach while intertwining the problem definition and analysis steps as in the incrementalist approach. In
practice, this means that the collective understanding of the problem and potential solutions must co-evolve. As Rittel
puts it, ``The systems-approach `of the first generation' [i.e., classical analytics] is inadequate for dealing with
wicked-problems. Approaches of the `second generation' should be based on a model of planning as an argumentative
process in the course of which an image of the problem and of the solution emerges gradually among the participants, as
a product of incessant judgment, subjected to critical argument'' \cite{rittel_dilemmas_1973}.

@@

This follows partially from the previous suggestion---problem definition and analysis are combined precisely because the
problem definition depends on the nature of proposed solutions. But this advice deserves emphasis for another reason:
consensus is easier to build around concrete proposals. While debate rages in the abstract, groups are bound to disagree
due to their conflicting values and priorities, but Linblom writes encouragingly about ``the ease with which individuals
of different ideologies often can agree on concrete policy'' in an example about congressional compromise; he goes on to
say, ``Labor mediators report a similar phenomenon: the contestants cannot agree on criteria for settling their disputes
but can agree on specific proposals. Similarly, when one administrator's objective turns out to be another's means, they
often can agree on policy'' \cite{lindblom_muddling_1959}.

>>>>

The failures of the \ac{classical-method} and \ac{incrementalism} teach that we must be able to think strategically
while respecting the non-linear nature of \acp{wicked-prob}. One way to use the high level, holistic view of the
classical approach while intertwining the problem definition and analysis steps as in the incrementalist approach. In
practice, this means that the collective understanding of the problem and potential solutions must co-evolve. As Rittel
puts it, ``The systems-approach `of the first generation' [i.e., classical analysis] is inadequate for dealing with
wicked-problems. Approaches of the `second generation' should be based on a model of planning as an argumentative
process in the course of which an image of the problem and of the solution emerges gradually among the participants, as
a product of incessant judgment, subjected to critical argument'' \cite{rittel_dilemmas_1973}.

@@

This follows partially from the previous suggestion---problem definition and analysis are combined precisely because the
problem definition depends on the nature of proposed solutions. But this advice deserves emphasis for another reason:
consensus is easier to build around concrete proposals. While debate rages in the abstract, groups are bound to disagree
due to their conflicting values and priorities, but Lindblom writes encouragingly about ``the ease with which individuals
of different ideologies often can agree on concrete policy'' in an example about congressional compromise; he goes on to
say, ``Labor mediators report a similar phenomenon: the contestants cannot agree on criteria for settling their disputes
but can agree on specific proposals. Similarly, when one administrator's objective turns out to be another's means, they
often can agree on policy'' \cite{lindblom_muddling_1959}.

### 2020-11-23 Thesis-Chapter5-ThreatModel.tex
Clean up glossary
Spelling and editorial improvements

To save space, the tables use shorthand for law enforcement (LE), law enforcement officer (LEO), law enforcement agency
(LEA), Device ID (DID), Device Seal Key (DSK), \acf{DOS}

@@

\threattablebegin{Tampering}{tampering}{Tamper}
  T1 & Defensive user corrupts DSK, DID, or token
  & LE attempts at perform \ac{EA} will fail due to corrupt data. Requires compromise of the Secure Enclave or advanced
    hardware manipulation tools.
  & Since these values are known at manufacturing time and never change, they can be made read-only. (The symmetric key
    variant of \ldawmsr requires updating the symmetric equivalent of the public DSK, and is therefore more vulnerable
    to this threat.)
  \\ \hline
  T2 & Defensive user destroys the \ac{EA} hardware interface
  & LE attempts to perform \ac{EA} will fail due to damaged hardware.
  & Manufacture the device to require destructive disassembly to access the interface. This increases the cost
    associated with deploying and using the system.
  \\ \hline
  T3 & LE insider tampers with DID before digital signature
  & LEA receives an unauthorized device's unlock artifacts. Can be combined with other attacks to compromise a device.
  & Randomize DIDs to make them unguessable, meaning a target device must be in possession in order for the attacker to
    know what DID to use. Input to the vendor's \ac{HSM} could also include data from the legal documents, such as
    device S/N; the \ac{HSM} would append the S/N to the DID and hash the result, verifying a match with the same stored
    hash before returning the unlock artifacts.
  \\ \hline
  T4 & Attacker tampers with hardware transparency mechanism
  & The device user is unaware that the device was unlocked. Highly implementation dependent, but certainly requires
    advanced hardware manipulation tools.
  & Manufacture the device to require destructive disassembly to access the interface.
\threattableend

@@

\paragraph*{Economic impact} Economic costs to the vendor fall into three categories---implementing the design, running
the system, and absorbing lost business. Even once a proposal like \ldawmsr was refined enough to be implemented, the
vendor would need to update hardware designs and create the software systems to get it started. Once in use, the vendor
would face considerable costs in running it securely, considering the supreme importance of transporting and storing the
unlock artifacts safely. Finally, the vendor would certainly lose privacy-conscious customers to vendors not under the
system, whether small businesses, foreign competitors, or resellers offering pre-\ac{EA} devices. The scale of these
costs relative to device manufacturer's naturally large size is unclear. Costs to the government include the \ac{FBI}
operation and a necessary scale-up in training and resources available to state and local law enforcement. Government
reimbursement for vendor compliance costs has precedent in laws like \acs{CALEA}, and could be applied to a \ac{DAR}
\ac{EA} regulation as well \cite{edwards_hr4922_1994}.

>>>>

To save space, the tables use shorthand for \acf{LE}, \acf{LEO}, \acf{LEA}, \acf{DID}, \acf{DSK}, and \acf{DOS}.

@@

\threattablebegin{Tampering}{tampering}{Tamper}
  T1 & Platform abuser corrupts DSK, DID, or token
  & LE attempts at perform \ac{EA} will fail due to corrupt data. Requires compromise of the Secure Enclave or advanced
    hardware manipulation tools.
  & Since these values are known at manufacturing time and never change, they can be made read-only. (The symmetric key
    variant of \ldawmsr requires updating the symmetric equivalent of the public DSK, and is therefore more vulnerable
    to this threat.)
  \\ \hline
  T2 & Platform abuser destroys the \ac{EA} hardware interface
  & LE attempts to perform \ac{EA} will fail due to damaged hardware.
  & Manufacture the device to require destructive disassembly to access the interface. This increases the cost
    associated with deploying and using the system.
  \\ \hline
  T3 & LE insider tampers with DID before digital signature
  & LEA receives an unauthorized device's unlock artifacts. Can be combined with other attacks to compromise a device.
  & Randomize DIDs to make them unguessable, meaning a target device must be in possession in order for the attacker to
    know what DID to use. Input to the vendor's \ac{HSM} could also include data from the legal documents, such as
    device S/N; the \ac{HSM} would append the S/N to the DID and hash the result, verifying a match with the same stored
    hash before returning the unlock artifacts.
  \\ \hline
  T4 & Attacker tampers with hardware transparency mechanism
  & The device user is unaware that the device was unlocked. Highly implementation dependent, but certainly requires
    advanced hardware manipulation tools.
  & Manufacture the device to require destructive disassembly to access the interface.
\threattableend

@@

\paragraph*{Economic impact} Economic costs to the vendor fall into three categories---implementing the design, running
the system, and absorbing lost business. Even once a proposal like \ldawmsr was refined enough to be implemented, the
vendor would need to update hardware designs and create the software systems to get it started. Once in use, the vendor
would face considerable costs in running it securely, considering the supreme importance of transporting and storing the
unlock artifacts safely. Finally, the vendor would certainly lose privacy-conscious customers to vendors not under the
system, whether small businesses, foreign competitors, or resellers offering pre-\ac{EA} devices. The scale of these
costs relative to device manufacturer's naturally large size is unclear. Costs to the government include the \ac{FBI}
laboratory cost and a necessary scale-up in training and resources available to state and local law enforcement.
Government reimbursement for vendor compliance costs has precedent in laws like \acs{CALEA}, and could be applied to a
\ac{DAR} \ac{EA} regulation as well \cite{edwards_hr4922_1994}.

### 2020-11-23 Thesis-Chapter6-PathsAhead.tex
Spelling and editorial improvements

Regarding the arguments analyzed in \mychap{chap-arguments}, while neither side of the \ac{EA} debate fully incorporates
the whole picture, the anti-\ac{EA} argument is currently stronger. Encryption poses a present danger to public safety
in certain respects. The bogeymen of terrorists and criminals using encryption to evade the law are real---in the case
of those peddling and viewing \acl{CSAM}, shockingly so \cite{keller_internet_2019}. However, the argument that
\ac{encryption} is creating circumstances that are to blame for \acl{LE} failure is not currently strong enough to
justify an \ac{EA} mandate. For \ac{DIM}, the primary issues are poor coordination and training \cite{carter_2018}; for
\ac{DAR}, commercial \ac{lawful-hacking} tools currently have the upper hand on mobile phone manufacturers, meaning
\acl{LE} can get into almost any mobile device they acquire \cite{koepke_2020}. Strong \ac{encryption} plays a key role
in protecting security and privacy, public goods that other technological developments often fail to provide.

@@

This thesis represents an ``OO'' in the \ac{OODAloop} (see \myfig{fig-policy-ooda-loop}). It is observation and
orientation, framed around the debate and a specific proposal. Observation and orientation are just the beginning of the
process; next it is time to decide and act. How should technology makers respond?

The technical community needs to accept and deal with the enormous role technology is playing in current policy issues,
such as encryption, automation, and misinformation. Government sometimes suggests that the tech world is is not trying
hard enough, and that ``nerding harder'' will result in solutions. The tech world is absolutely right that nerding
harder will not work on these issues---because they are \acp{wicked-prob}---but that does not mean they cannot deal with
them. Tech is used to addressing tame problems, but it is time that they turn their massive operations and intellectual
capital towards strategies designed for \acp{wicked-prob}. This means including people of diverse experiences,
backgrounds, and values in decision making processes that steer investment and design decisions. It means creating
products and business models that are responsive to evolving conditions. It means considering the ways their creations
will be abused.

@@

\section{Paths Forward: Policy Makers}

Government has not been silent on the issue of encryption and lawful access. The dynamics created by the executive,
legislative, and judicial branches ensures that the regulatory environment will evolve with changing federal policies,
new laws, and rulings that refine implementation existing law. The judiciary will indeed shape encryption policy through
its rulings, and is correct to be reevaluating current Fourth Amendment jurisprudence in light of rapidly changing
circumstances. Still, the primary actors capable of consciously taking part in strategic policymaking will be the
executive and legislative branches. How should these policymakers respond?

Government must strategically approach encryption policy as a \ac{wicked-prob}. Industry has failed at analyzing the
problem away, and government has failed at muddling the problem away. It is true that incrementalism has safely avoided
dangerous outcomes such as blanket \ac{EA} mandates, but it has also resulted in a research-stifling high-tension debate
and circumstances of poor utilization of current capabilities mixed with unregulated use of invasive \ac{lawful-hacking}
tools. Policymakers and law enforcement agencies need to be forthcoming with the real challenges they face. They are
important contributors to the diverse groups that can iteratively refine the problem definition and possible solutions.
Just like the technical side, the policy side of the solutions must be responsive to evolving conditions. Most
importantly, any initiative needs follow through, since a single \ac{OODAloop} iteration will not be enough to tame this
issue.

@@

Law enforcement is often ill-equiped to handle digital evidence and coordinate with servive providers using existing
access mechanisms \cite{carter_2018}. Justice Department task forces to combat \acl{CSAM} are underfunded and
understaffed \cite{keller_internet_2019}. Improving the effectiveness of existing capabilities and programs carries no
risk and guaranteed results.

@@

The goal of this thesis is to clarify encryption and \ac{EA} as a \ac{wicked-prob}, to break down the debate, and to
evaluate the security of current \ac{EA} proposals. I aim to move the discussion forward with the ideas of how to tackle
\acp{wicked-prob}, additions to the traditional cryptosystem threat model, and security analysis of a modern proposal.
This study is observation and orientation; it is up to technologists, policymakers, and common people to decide and
act. Contending with these issues is difficult on many levels, but the future of security, safety, and privacy depend on
coming together to find a good answer where no perfect answer exists.

>>>>

Regarding the arguments analyzed in \mychap{chap-arguments}, while neither side of the \ac{EA} debate fully incorporates
the whole picture, the anti-\ac{EA} argument is currently stronger. Encryption poses a present danger to public safety
in certain respects. The bogeymen of terrorists and criminals using encryption to evade the law are real---in the case
of those peddling and viewing \acl{CSAM}, shockingly so \cite{keller_internet_2019}. However, the argument that
\ac{encryption} is creating circumstances that are to blame for \acl{LE} failure is not currently strong enough to
justify an \ac{EA} mandate. For \ac{DIM}, the primary issues are poor coordination and training \cite{carter_2018}; for
\ac{DAR}, commercial \ac{lawful-hacking} tools currently have the upper hand on mobile phone manufacturers, meaning
\acl{LE} can get into almost any mobile device they acquire \cite{koepke_2020}. Additionally, strong \ac{encryption}
plays a key role in protecting security and privacy, public goods that other technological developments often fail to
provide.

@@

The technical community needs to accept and deal with the enormous role technology is playing in current policy issues,
such as encryption, automation, and misinformation. Government sometimes suggests that the tech world is is not trying
hard enough, and that ``nerding harder'' will result in solutions. The tech world is absolutely right that nerding
harder will not work on these issues---because they are \acp{wicked-prob}---but that does not mean they cannot deal with
them. Tech is used to addressing tame problems, but it is time that they turn their massive financial and intellectual
capital towards strategies designed for \acp{wicked-prob}. This means including people of diverse experiences,
backgrounds, and values in decision making processes that steer investment and design decisions. It means creating
products and business models that are responsive to evolving conditions. It means considering the ways their creations
will be abused.

@@

\section{Paths Forward: Policymakers}

Government has not been silent on the issue of encryption and lawful access. The dynamics created by the executive,
legislative, and judicial branches ensures that the regulatory environment will evolve with changing federal policies,
new laws, and rulings that refine implementation existing law. The judiciary will indeed shape encryption policy through
its rulings, and is correct to be reevaluating current Fourth Amendment jurisprudence in light of rapidly changing
circumstances. Still, the primary actors capable of consciously taking part in strategic policymaking will be the
executive and legislative branches.

Executive and legislative policymakers must strategically approach encryption policy as a \ac{wicked-prob}. Industry has
failed at analyzing the problem away, and government has failed at muddling the problem away. It is true that
incrementalism has safely avoided dangerous outcomes such as blanket \ac{EA} mandates, but it has also resulted in a
research-stifling high-tension debate and circumstances of poor utilization of current capabilities mixed with
unregulated use of invasive \ac{lawful-hacking} tools. Policymakers and law enforcement agencies need to be forthcoming
with the real challenges they face. They are important contributors to the diverse groups that can iteratively refine
the problem definition and possible solutions. Just like the technical side, the policy side of the solutions must be
responsive to evolving conditions. Most importantly, any initiative needs follow through, since a single \ac{OODAloop}
iteration will not be enough to tame this issue.

@@

Law enforcement is often ill-equipped to handle digital evidence and coordinate with service providers using existing
access mechanisms \cite{carter_2018}. Justice Department task forces to combat \acl{CSAM} are underfunded and
understaffed \cite{keller_internet_2019}. Improving the effectiveness of existing capabilities and programs carries no
risk and guaranteed results.

@@

The goal of this thesis is to clarify encryption and \ac{EA} as a \ac{wicked-prob}, to break down the debate, and to
evaluate the security of current \ac{EA} proposals. I aim to move the discussion forward with the ideas of how to tackle
\acp{wicked-prob}, additions to the traditional cryptosystem threat model, and security analysis of a modern proposal.
In the language of the \ac{OODAloop}, this study is observation and orientation; it is up to technologists,
policymakers, and common people to decide and act. Contending with these issues is difficult on many levels, but the
future of security, safety, and privacy depend on coming together to find a good answer where no perfect answer exists.
